\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization

### 2010-04-20 honors_thesis/ppsn_paper/paper.tex
fixed some formatting in equations and cut a line
combined equations for compactness

@@

In order to incorporate the determination of which case occurs into the
position and velocity update equations, we introduce an indicator function
$\indic_{t+1}^{\casegen}$ for each case $\casegen \in \caseset$.  When
$\casegen$ corresponds to the case taken by PSO,
$\indic_{t+1}^{\casegen}$ evaluates to 1; otherwise it evaluates to 0.  We can
then sum over all of the cases, and the indicator function will make all of the
terms drop to zero except for the case that actually occurs.  For example, the
indicator function for the specific case $\casexn$ can be written as follows:

\begin{align}
  \nonumber
	\indic_{t+1}^{\casexn} & (\ofunc ( \ppos_{t} ) ,\ofunc(\leftn_{t}),
	\ofunc(\rightn_{t}) ,\ofunc(\pbest_{t-1}) ,\ofunc(\nbest_{t-1}))= \\
  \label{eq:deficasexn}
	&\begin{cases}
	   1 & \text{if} \ [ \ofunc(\ppos_{t}) < \ofunc(\pbest_{t-1}) ]
	    \wedge [ \ofunc(\nbest_{t-1}) < \ofunc(\ppos_{t}) ] \\
	   &\quad \wedge \quad [ \ofunc(\nbest_{t-1}) < \ofunc(\leftn_{t}) ]
	    \wedge [ \ofunc(\nbest_{t-1}) < \ofunc(\rightn_{t}) ] \\
	   0 & \text{otherwise}
	\end{cases}
\end{align}

For each case $\casegen \in \caseset$, there is also a corresponding
velocity update function $\specvel_{t+1}^{\casegen}$.  When the case is
known, the specific values of $\pbest_t$ and $\nbest_t$ may be substituted
directly into~\eqref{eq:velupdate}.  For example, in case $\casexn$,
$\pbest_{t}=\ppos_{t}$, as \pbest was updated by the particle's current
position, and $\nbest_{t}=\nbest_{t-1}$, as $\nbest$ was not updated at
iteration $t$:
\begin{align}
\nonumber
	\specvel_{t+1}^{\casexn} & (\pvel_t, \ppos_{t}, \leftn_{t}, \rightn_{t},
	\pbest_{t-1}, \nbest_{t-1}, \pURand_{t}, \nURand_{t}) \\
\label{eq:defvcasexn}
		&= \constriction \bigl[ \pvel_{t} +
			\pcoeff\pURand_{t}\otimes(\ppos_{t} - \ppos_{t})
			+ \ncoeff\nURand_{t}\otimes(\nbest_{t-1} -
			\ppos_{t}) \bigr]
\end{align}

In the same way we can create notation for the position update function by
substituting into~\eqref{eq:posupdate}.  For compactness, we will drop the
parameters to $\specvel_{t+1}^{\casegen}$ since they can be inferred from the
subscripts.
\begin{align}
\label{eq:defpcasegen}
	\specpos_{t+1}^{\casegen} & (\ppos_{t}, \pvel_{t}, \leftn_{t},
	\rightn_{t} ,\pbest_{t-1} ,\nbest_{t-1}, \pURand_{t}, \nURand_{t})
	 = \ppos_{t} + \specvel_{t+1}^{\casegen}
\end{align}

With this notation we can re-write the original PSO velocity
equation~\eqref{eq:velupdate}, introducing our sum over cases with the
indicator functions.  Again, we represent the indicator functions and velocity
functions without the parameters for compactness.  The equation becomes:
\begin{align}
\nonumber
	\pvel_{t+1} &=
		\constriction \bigl[ \pvel_t
			+ \pcoeff\pURand_{t}\otimes(\pbest_{t} - \ppos_{t})
			+ \ncoeff\nURand_{t}\otimes(\nbest_{t} -
			\ppos_{t}) \bigr] \\
\nonumber
	&= \sum_{c \in \caseset} \indic_{t+1}^{c} \ \constriction \bigl[ \pvel_t
			+ \pcoeff\pURand_{t}\otimes(\pbest_{t} - \ppos_{t})
			+ \ncoeff\nURand_{t}\otimes(\nbest_{t} -
			\ppos_{t}) \bigr]  \\
\label{eq:vel2update}
	&= \sum_{c \in \caseset} \ \indic_{t+1}^{c} \ \specvel_{t+1}^{c} 
\end{align}

Similarly, the position update equation~\eqref{eq:posupdate} becomes:
\begin{align}
\label{eq:pos2update}
	\ppos_{t+1} &= \ppos_{t} + \pvel_{t+1}
	= \sum_{c \in \caseset} \ \indic_{t+1}^{c} \ \specpos_{t+1}^{c} 
\end{align}

Returning our attention to the computation of $\ppos_{t+1}$ in
\eqref{eq:pos2update} and writing it with the parameters which were omitted
above, we obtain:
\begin{align}
\nonumber
  \ppos_{t+1} &= \sum_{c \in \caseset} \bigl[
	\indic_{t+1}^{c}(\ofunc ( \ppos_{t} ) ,\ofunc(\leftn_{t}),
	\ofunc(\rightn_{t}) ,\ofunc(\pbest_{t-1}) ,\ofunc(\nbest_{t-1})) \\
\label{eq:val2updatelong}
	& \quad \quad \quad \specpos_{t+1}^{c}(\ppos_{t},\pvel_{t},\leftn_{t},\rightn_{t},
	\pbest_{t-1},\nbest_{t-1},\pURand_{t}, \nURand_{t}) \bigr]
\end{align}

>>>>

% Mathy help
\newcommand{\parens}[1]{\!\left(#1\right)}

@@

In order to incorporate the determination of which case occurs into the
position and velocity update equations, we introduce an indicator function
$\indic_{t+1}^{\casegen}$ for each case $\casegen \in \caseset$.  When
$\casegen$ corresponds to the case taken by PSO,
$\indic_{t+1}^{\casegen}$ evaluates to 1; otherwise it evaluates to 0.  We can
then sum over all of the cases, and the indicator function will make all of the
terms drop to zero except for the case that actually occurs.  For example, the
indicator function for the specific case $\casexn$ can be written as follows:
$f(x)$
\begin{align}
  \nonumber
	&\indic_{t+1}^{\casexn}\parens{
	  \ofunc\parens{\ppos_{t}},
	  \ofunc\parens{\leftn_{t}},
	  \ofunc\parens{\rightn_{t}},
	  \ofunc\parens{\pbest_{t-1}},
	  \ofunc\parens{\nbest_{t-1}}
	  } \\
  &\hspace{15pt}= \label{eq:deficasexn}
	\begin{cases}
	  1 & \text{if}~
		\ofunc\parens{\ppos_{t}} < \ofunc\parens{\pbest_{t-1}}
	    \,\text{and}\,
		\ofunc\parens{\nbest_{t-1}} < \min\parens{
		  \ofunc\parens{\ppos_{t}},
		  \ofunc\parens{\leftn_{t}},
		  \ofunc\parens{\rightn_{t}}
		} \\
	   0 & \text{otherwise}
	\end{cases}
\end{align}

For each case $\casegen \in \caseset$, there is also a corresponding
velocity update function $\specvel_{t+1}^{\casegen}$.  When the case is
known, the specific values of $\pbest_t$ and $\nbest_t$ may be substituted
directly into~\eqref{eq:velupdate}.  For example, in case $\casexn$,
$\pbest_{t}=\ppos_{t}$, as \pbest was updated by the particle's current
position, and $\nbest_{t}=\nbest_{t-1}$, as $\nbest$ was not updated at
iteration $t$:
\begin{align}
\nonumber
	\specvel_{t+1}^{\casexn} & \parens{\pvel_t, \ppos_{t}, \leftn_{t}, \rightn_{t},
	\pbest_{t-1}, \nbest_{t-1}, \pURand_{t}, \nURand_{t}} \\
\label{eq:defvcasexn}
		&= \constriction \left[ \pvel_{t} +
			\pcoeff\pURand_{t}\otimes\parens{\ppos_{t} - \ppos_{t}}
			+ \ncoeff\nURand_{t}\otimes\parens{\nbest_{t-1} -
			\ppos_{t}} \right]
\end{align}

In the same way we can create notation for the position update function by
substituting into~\eqref{eq:posupdate}.  For compactness, we will drop the
parameters to $\specvel_{t+1}^{\casegen}$ since they can be inferred from the
subscripts.
\begin{align}
\label{eq:defpcasegen}
	\specpos_{t+1}^{\casegen} & \parens{\ppos_{t}, \pvel_{t}, \leftn_{t},
	\rightn_{t} ,\pbest_{t-1} ,\nbest_{t-1}, \pURand_{t}, \nURand_{t}}
	 = \ppos_{t} + \specvel_{t+1}^{\casegen}
\end{align}

With this notation we can re-write the original PSO velocity
equation~\eqref{eq:velupdate}, introducing our sum over cases with the
indicator functions.  Again, we represent the indicator functions and velocity
functions without the parameters for compactness.
The velocity update from~\eqref{eq:velupdate} becomes:
\begin{align}
\label{eq:vel2update}
%\nonumber
	\pvel_{t+1} 
%		\constriction \bigl[ \pvel_t
%			+ \pcoeff\pURand_{t}\otimes\parens{\pbest_{t} - \ppos_{t}}
%			+ \ncoeff\nURand_{t}\otimes\parens{\nbest_{t} -
%			\ppos_{t}} \bigr] \\
%\nonumber
%	&= \sum_{c \in \caseset} \indic_{t+1}^{c} \ \constriction \bigl[ \pvel_t
%			+ \pcoeff\pURand_{t}\otimes\parens{\pbest_{t} - \ppos_{t}}
%			+ \ncoeff\nURand_{t}\otimes\parens{\nbest_{t} -
%			\ppos_{t}} \bigr]  \\
	&= \sum_{c \in \caseset} \ \indic_{t+1}^{c} \ \specvel_{t+1}^{c} 
\end{align}

Similarly, the position update from~\eqref{eq:posupdate}
can also be rewritten using the indicator functions, now with the parameters reinserted:
\begin{align}
\nonumber
  \ppos_{t+1} &= \sum_{c \in \caseset} \bigl[
	\indic_{t+1}^{c}\parens{
	  \ofunc\parens{\ppos_{t}},
	  \ofunc\parens{\leftn_{t}},
	  \ofunc\parens{\rightn_{t}},
	  \ofunc\parens{\pbest_{t-1}},
	  \ofunc\parens{\nbest_{t-1}}
	  } \\
\label{eq:val2updatelong}
	& \quad \quad \quad \specpos_{t+1}^{c}\parens{
	  \ppos_{t},
	  \pvel_{t},
	  \leftn_{t},
	  \rightn_{t},
	  \pbest_{t-1},
	  \nbest_{t-1},
	  \pURand_{t},
	  \nURand_{t}
	  }\bigr]
\end{align}
